---
title: "Relationship Between Political Alignment and Self-Identification as an Environmentalist"
author: "Travis Christy"
date: "12/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(broom)
```

### Introduction:
This report will examine the relationship between self-identification as an environmentalist and political alignment as either a conservative or liberal. The data was collected from an anonymous online survey in August of 2017 regarding influences on charitable giving for conservation (see citation). This report first organizes data into tables then uses the chi-squared test on the contingency table to test for variable independence.  

  
  
  
---    
  
  
```{r}
# reading in data and creating simplified version
conservation_survey_simple <- read_csv(here("data", "conservation_survey.csv")) %>% 
  clean_names() %>% 
  select(env_id, pol_or)

# organizing by political allignment into only two values representing conservative (pol_or = 1-3) or liberal (pol_or = 5-7)
survey_binary_party <- conservation_survey_simple %>% 
  filter(pol_or == c(1, 2, 3, 5, 6, 7)) %>% 
  mutate(pol_align = case_when(pol_or <= 3 ~ "Conservative",
                               pol_or >= 5 ~ "Liberal"))
```

#### Table 1. Survey responses to env_id question aka "I consider myself an environmentalist".

```{r}
# creating table for answer to env_id question by political alignment
q1_counts <- survey_binary_party %>% 
  tabyl(pol_align, env_id) %>% 
  rename("Political Alignment" = "pol_align", "Yes" = "1", "No" = "2", "Unsure" = "3")

# create proportions table
q1_proportions <- q1_counts %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting(digits = 2) %>% 
  adorn_ns()

# display finalized table
kbl(q1_proportions) %>% 
    kable_material(full_width = F)
```
  
--- 
  
#### Chi-squared test for variable independence between self-identification as an environmentalist and political alignment.

```{r}
# convert data to contingency table
q1_ct <- q1_counts %>% 
  column_to_rownames(var = "Political Alignment")

# run chi-square test
q1_x2 <- chisq.test(q1_ct)

# convert to tidy data for in line referencing
q1_tidy <- tidy(q1_x2)
```

Political alignment and self-identification as an environmentalist are not independent (there is a significant association between political alignment and response to the env_id question of the survey) :  $\chi$^2^(`r q1_tidy$parameter`) = `r round(q1_tidy$statistic,2)`, *p* = `r round(q1_tidy$p.value,3)`  
  
  The majority of conservatives do not identify as environmentalists and the majority of liberals do identify as environmentalists. 
  
---
  
**CITATION**: Andrews Forest LTER Site and C. Batavia. 2019. Influences on charitable giving for conservation: Online survey data of 1,331 respondents across the US, August 2017 ver 3. Environmental Data Initiative. [URL LINK](https://doi.org/10.6073/pasta/926e6270e324a1322a900da14d38b96c)   




















