---
title: "a6_task4_travis_christy"
author: "Travis Christy"
date: "12/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(broom)
library(ggbeeswarm)
library(effsize)
library(ggpubr)
library(dplyr)
library(plotly)
```

```{r}
# reading in data and creating simplified version
mono_tc <- read_csv(here("data", "mono_lake_annual_levels.csv"), skip = 5) %>% 
  clean_names()
```




```{r,}


fig1 <- plot_ly(
  type = "scatter",
  x = mono_tc$year, 
  y = mono_tc$lake_level_feet_above_sea_level,
  name = 'Lake Level (ft above sea-level)',
  mode = "lines",
  line = list(
        color = '#17BECF'
  )) 
fig1 <- fig1 %>%
  add_trace(
    type = "scatter",
    x = mono_tc$year, 
    y = mono_tc$stable_level,
    name = 'Management Level',
    mode = "lines",
    line = list(
        color = '#7F7F7F'
  )) 
fig1 <- fig1 %>%
  add_trace(
    type = "scatter",
    x = mono_tc$year, 
    y = 6377,
    name = 'Predator Land Bridge Level',
    mode = "lines",
    line = list(
        color = 'red'
  )) 
fig1 <- fig1 %>%
  add_trace(
    type = "scatter",
    x = mono_tc$year, 
    y = 6360,
    name = 'Salinity 120/g/L level',
    mode = "lines",
    line = list(
        color = 'darkgoldenrod'
  )) 
fig1 <- fig1 %>%
  add_trace(
    x = 1994,
    y = 6374.6,
    name = "Decision 1631 sets 'Management Label'",
    marker = list(
      color = 'rgb(17, 157, 255)',
      size = 10,
      line = list(
        color = 'rgb(231, 99, 250)',
        width = 2
      )
    ),
    showlegend = T
  )
fig1 <- fig1 %>%
  layout(
    title = "Mono Lake Levels 1850-2017 (CHRISTY 2020)",
    yaxis = list(
      title = 'Water-level in feed above sea-level'),
    xaxis = list(
        title = 'Year',
        type = 'date',
        tickformat = "%Y"
  ))

fig1
```
**Figure 1:** The Mono Lake Management Level was created in 1994 due to the implications for nesting birds (eared grebes and California gulls) that rely on the safety of internal nesting islands for protection from predators and food. When the water level drops below 6377 ft above sea-level, predators can cross the newly formed land bridge and prey on the nesting birds and offspring. When the water levels drop below 6360 ft above sea-level, salinity levels begin to negatively affect biological food sources for the birds. The management level restricts LADW water extraction.
  
---
  
  
**CITATION**: 2020 Mono Lake Committee [URL LINK](https://www.monolake.org/learn/aboutmonolake/savingmonolake/d1631/)   